---
title: "Spanish Analysis"
author: "Dan Kats and Mondira Ray"
date: "2024-10-30"
output: html_document
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries}
library(MASS)
library(tidyr)
library(dplyr)
library(ggplot2)
library(viridis)
library(readxl)
library(grid)
```

# Data loading and processing
Files are organized as folders named by lingust/evaluator (e.g., "Linguist 1_DK"). Inside each of those folders, there are folders for each pair of translations with a DOCX containing both (human and GPT) translations and their error annotations and two XLSX scorecards with one for each of the translations.
```{r Data loading}
# Get the linguists
files <- list.files()
# Filter to those that start with "Linguist"
linguists <- grep("^Linguist", files, value = TRUE)
linguists <- linguists[file.info(linguists)$isdir]
print(linguists)

# Get the key of which translations were GPT vs human (HT = Human Translation; GPT = GPT Translation)
translation_key <- read_excel('Translation Units.xlsx')
translation_key$unit <- sub("^[^-]*-", "", translation_key$ID)
translation_key$HT <- gsub("[^0-9]", "", translation_key$HT)
translation_key$GPT <- gsub("[^0-9]", "", translation_key$GPT)

# Load in which translations were preferred by the evaluators
preferences <- read_excel('preferences.xlsx') |>
	pivot_longer(cols = contains("prefer", ignore.case=TRUE), names_to = "preference", values_to = "x", values_drop_na = TRUE) |>
	select(-"x") |>
	mutate(preference_version = gsub("[^0-9]", "", preference)) |>
	mutate(preference_version = gsub("^$", 0, preference_version)) |>
	mutate(preference_version = as.integer(preference_version)) |>
	mutate(preference_strength = case_when(
		grepl("^Prefer", preference) ~ 1,
		grepl("^Strongly", preference) ~ 2,
		TRUE ~ 0
	)) |>
	mutate(preference_strength = as.integer(preference_strength)) |>
	select(-"preference")

# Get contents of the working directory
files <- list.files(recursive = TRUE)
# Filter to those that start with "Linguist"
translations <- data.frame(file = grep("^Linguist.*\\.xlsx$", files, value = TRUE))

translations$unit <- sapply(translations$file, function(x) strsplit(basename(x), "_")[[1]][1])
translations$linguist <- sapply(translations$file, function(x) strsplit(x, "/")[[1]][1])
translations$translation_id <- sapply(translations$file, function(x) strsplit(basename(x), "_")[[1]][2])
translations$translation_id <- gsub("[^0-9]", "", translations$translation_id)

translations$score_table <- lapply(translations$file, function(x) {
	out <- read_excel(x)
	out <- out[2:nrow(out), 3:ncol(out)]
	names(out) <- unlist(out[1,])
	names(out)[1] <- out[3,1]
	out <- out[-c(1,2,3),]
	out <- out |> mutate(across(-1, ~as.numeric(.))) |> 
		mutate(across(-1, ~replace_na(.,0)))
	return(out)
})

# Merge the evaluator preferences with the translation table
translations <- left_join(translations, preferences |> select(-'why'), by=c("linguist", "unit"))
translations <- translations |>
	mutate(preference = ifelse(preference_version == translation_id, 1, -1) * preference_strength) |>
	select(-c(preference_version, preference_strength))

# Merge the translation key with the translation table
translations <- left_join(translations, translation_key[c('unit', 'HT', 'GPT', 'HT EWC', 'GPT EWC')], by="unit")
translations$label <- ifelse(translations$translation_id == translations$HT, "Human", "GPT")
translations$EWC <- ifelse(translations$translation_id == translations$HT, translations$`HT EWC`, translations$`GPT EWC`)
translations <- translations |> select(-c('HT', 'GPT', 'HT EWC', 'GPT EWC'))
```

# MQM Scoring
```{r MQM}
### Parameters
# Reference Word Count: an arbitrary number of words in a hypothetical reference evaluation text. Implementers use this uniform word count to compare results across different projects. The RWC is often set at 1000.
RWC <- 1000
# Maximum Score Value: also an arbitrary value designed to manipulate the Quality Score in order to shift its value into a range which is easier to understand. It converts the score to a percentage-like value
MSV <- 100
# Defined Passing Threshold: the score that defines the Pass/Fail limit. Scoring methods without calibration typically use values such as 0.99 OR 99 – depending on the scale used. If scoring with calibration is used, the implementer can define any number that is perceived to be visually meaningful, such as 95 or 90.
DPT <- 99
# Defined Passing Interval: the interval between the Maximum Score Value and the Passing Threshold
DPI <- MSV-DPT
# Error Type Weights
ETWs <- rep(1, length(translations$score_table[[1]]['Error Types'][[1]]))
names(ETWs) <- translations$score_table[[1]]['Error Types'][[1]]
# Severity Penalty Multiplier
SPMs <- c(0,1,5,25)
names(SPMs) <- c('Neutral', 'Minor', 'Major', 'Critical')
# Acceptable Penalty Points for the Reference Word Count
APP <- DPI / MSV * RWC
# Scaling Factor: Parameter to scale the Acceptable Penalty Points for the reference word count across the Defined Passing Interval.
SF <- DPI/APP

## Calculations
# Count errors ("EC" = "Error Count")
translations$EC_neutral <- sapply(translations$score_table, function(x) sum(x[2]))
translations$EC_minor <- sapply(translations$score_table, function(x) sum(x[3]))
translations$EC_major <- sapply(translations$score_table, function(x) sum(x[4]))
translations$EC_critical <- sapply(translations$score_table, function(x) sum(x[5]))
translations$EC_significant <- translations$EC_major + translations$EC_critical
translations$EC_nonneutral <- translations$EC_minor + translations$EC_major + translations$EC_critical
translations$EC_total <- translations$EC_neutral + translations$EC_minor + translations$EC_major + translations$EC_critical
# Count error types
errortypes <- as.data.frame(sapply(translations$score_table, function(x) rowSums(x[,-1])))
rownames(errortypes) <- unlist(translations$score_table[[1]][,1])
translations$EC_wrongmedicalterm <- unlist(errortypes["Wrong medical term",])
translations$EC_mistranslation <- unlist(errortypes["Mistranslation",])
translations$EC_addition <- unlist(errortypes["Addition",])
translations$EC_omission <- unlist(errortypes["Omission",])
translations$EC_grammarspelling <- unlist(errortypes["Grammar, Spelling, Punctuation",])
translations$EC_register <- unlist(errortypes["Language register",])
translations$EC_awkward <- unlist(errortypes["Awkward style",])
translations$EC_insensitive <- unlist(errortypes["Culturally insensitive",])
# Penalty totals
translations$PT_neutral <- sapply(translations$score_table, function(x) sum(x[2] * ETWs))
translations$PT_minor <- sapply(translations$score_table, function(x) sum(x[3] * ETWs))
translations$PT_major <- sapply(translations$score_table, function(x) sum(x[4] * ETWs))
translations$PT_critical <- sapply(translations$score_table, function(x) sum(x[5] * ETWs))
# Absolute Penalty Total
translations <- translations |>
	mutate(APT = SPMs['Neutral'] * PT_neutral + SPMs['Minor'] * PT_minor + SPMs['Major'] * PT_major + SPMs['Critical'] * PT_critical)
# Per-Word Penalty Total
translations$PWPT <- translations$APT / translations$EWC
# Normed Penalty Total: Per-Word Error Penalty total relative to the Reference Word Count
translations$NPT <- translations$PWPT * RWC
# Raw Quality Score
translations$RQS <- 100 - 100 * translations$PWPT
# Calibrated Quality Score
translations$CQS <- 100 - translations$NPT * SF

translations |> 
	select(-score_table) |> 
	write.csv(file="translation_scores.csv")
```

# Graphing
```{r Graphing setup}
background_data <- data.frame(
	unit = unique(translations$unit),
	xmin = seq_along(unique(translations$unit)) - 0.5,
	xmax = seq_along(unique(translations$unit)) + 0.5,
	fill = rep(c("white", "gray95"), length.out = length(unique(translations$unit)))
)

col_gpt = "#21908CFF"
col_neutral = "#F1E51DFF"
col_human = "#472D7BFF"
```

```{r Bar plot of all ratings}
translations |>
	arrange(unit, label, CQS) |>
	mutate(order = interaction(unit, label, linguist, CQS), order = forcats::fct_inorder(order)) |>
	ggplot(aes(x = as.numeric(factor(unit)), y = CQS, fill = label, group = order)) +
		geom_rect(data = background_data, xmin = background_data$xmin, xmax = background_data$xmax, ymin = -Inf, ymax = Inf, fill = background_data$fill, inherit.aes = FALSE, alpha = 0.5) +
		geom_bar(stat = "identity", position = "dodge", color = "black") +
		geom_segment(inherit.aes = F, data = translations |> group_by(unit, label) |> summarize(mean_score = mean(CQS), .groups = 'drop') |>
					 	mutate(xmin = as.numeric(factor(unit)) - 0.2, xmax = as.numeric(factor(unit)) + 0.2),
					 aes(x = xmin-0.25, xend = xmax+0.25, y = mean_score, yend = mean_score, color = label), linewidth = 2) +
		labs(x = "Translation Unit", y = "MQM Score", fill = NULL, color = "Mean") +
		scale_fill_manual(values = c(col_gpt, col_human, col_neutral)) +
		scale_color_manual(values = c(col_gpt, col_human, col_neutral)) +
		scale_x_continuous(breaks = unique(as.numeric(factor(translations$unit)))) +
		theme_bw() +
		theme(
			panel.grid.major.x = element_blank(),
			panel.grid.minor.x = element_blank()
		) +
		guides(fill = guide_legend(order = 1), color = guide_legend(order = 2))
```
```{r Scatter plot of all error counts}
p <- translations |>
	ggplot(aes(x = as.numeric(factor(unit)), y = EC_nonneutral, fill = label)) +
		geom_rect(data = background_data, xmin = background_data$xmin, xmax = background_data$xmax, ymin = -Inf, ymax = Inf, fill = background_data$fill, inherit.aes = FALSE, alpha = 0.5) +
		geom_point(position = position_dodge(width = 0.8), shape = 21, size = 3, color = "black", alpha = 0.8) +
		labs(x = "Translation", y = "Number of Non-neutral Errors", fill = NULL, title = "Distribution of Non-neutral Errors by Translation") +
		scale_fill_manual(values = c(col_gpt, col_human, col_neutral)) +
		scale_x_continuous(breaks = unique(as.numeric(factor(translations$unit)))) +
		theme_bw() +
		theme(
			panel.grid.major.x = element_blank(),
			panel.grid.minor.x = element_blank(),
			plot.title = element_text(hjust = 0.5, size = 18),
			axis.title.x = element_text(size = 14),
			axis.title.y = element_text(size = 14),
			legend.text = element_text(size = 14)
		)
p
ggsave("error-distribution.png", plot = p, width = 10, height = 4, dpi = 1200)
```

```{r Scatter plot of all scores}
p <- translations |>
	ggplot(aes(x = as.numeric(factor(unit)), y = CQS, fill = label)) +
		geom_rect(data = background_data, xmin = background_data$xmin, xmax = background_data$xmax, ymin = -Inf, ymax = Inf, fill = background_data$fill, inherit.aes = FALSE, alpha = 0.5) +
		geom_point(position = position_dodge(width = 0.8), shape = 21, size = 3, color = "black", alpha = 0.8) +
		# geom_segment(inherit.aes = F, data = translations |> group_by(unit, label) |> summarize(mean_score = mean(CQS), .groups = 'drop') |>
		# 		  	mutate(xmin = as.numeric(factor(unit)) - 0.2, xmax = as.numeric(factor(unit)) + 0.2),
		# 			aes(x = xmin, xend = xmax, y = mean_score, yend = mean_score, color = label), linewidth = 1) +
		labs(x = "Translation", y = "MQM Score", fill = NULL, title = "Distribution of MQM Scores by Translation") +
		scale_fill_manual(values = c(col_gpt, col_human, col_neutral)) +
		scale_x_continuous(breaks = unique(as.numeric(factor(translations$unit)))) +
		theme_bw() +
		theme(
			panel.grid.major.x = element_blank(),
			panel.grid.minor.x = element_blank(),
			plot.title = element_text(hjust = 0.5, size = 18),
			axis.title.x = element_text(size = 14),
			axis.title.y = element_text(size = 14),
			legend.text = element_text(size = 14)
		)
p
ggsave("score-distribution.png", plot = p, width = 10, height = 4, dpi = 1200)
```
```{r Histogram of preferences, message=FALSE, warning=FALSE}
preference_labels = c("Strongly Prefer Human", "Prefer Human", "Neutral", "Prefer GPT", "Strongly Prefer GPT")
p <- translations |>
	filter(label == "GPT") |>
	select(preference) |>
	mutate(label = case_when(
		preference > 0 ~ "GPT",
		preference < 0 ~ "Human",
		TRUE ~ "Neither"
	)) |>
	mutate(preference = factor(preference_labels[preference - min(preference) + 1], levels = preference_labels)) |>
	ggplot(aes(x = preference, fill = label)) +
		geom_histogram(stat = "count") +
		labs(x = NULL, y = "Frequency") +
		scale_fill_manual(values = c(col_gpt, col_human, col_neutral)) +
		scale_y_continuous(breaks = function(x) pretty(x, n = 4, min.n = 1)) +
		theme_bw() +
		theme(legend.position = "none", text = element_text(size = 12))
p
ggsave("preference.png", plot = p, width = 7, height = 4, dpi = 1200)
```

```{r Scatter plot of all ratings}
p <- translations |>
	filter(label == "GPT") |> 
	ggplot(aes(x = as.numeric(factor(unit)) + as.numeric(factor(linguist)) * 0.2 - 0.4, y = preference)) +
		scale_x_continuous(breaks = unique(as.numeric(factor(translations$unit))),
				minor_breaks = seq(min(unique(as.numeric(factor(translations$unit)))), max(unique(as.numeric(factor(translations$unit)))), by = 1)+0.5,
				labels = unique(as.numeric(factor(translations$unit)))) +
		scale_y_continuous(breaks = seq(-2,2), labels = c("Strongly prefer human", "Prefer human", "Neutral", "Prefer GPT", "Strongly prefer GPT")) +
		labs(x = "Translation Unit", y = NULL, color = NULL) +
		theme_bw()

p <- p + annotation_custom(
		rasterGrob(as.raster(matrix(colorRampPalette(c(col_gpt, "white", col_human))(100), ncol=1)), width = unit(1, "npc"), height = unit(1, "npc")),
		ymin = -Inf, ymax = Inf, xmin = -Inf, xmax = Inf
	)

p <- p + theme(
		panel.background = element_rect(fill = "transparent", colour = NA),
		plot.background = element_rect(fill = "transparent", colour = NA)
	)

p <- p + geom_rect(
	data = translations |> 
		filter(label == "GPT") |> 
		group_by(unit) |> 
		reframe(
			ymin = min(preference),
			ymax = max(preference),
			x = as.numeric(factor(unit))
		),
	aes(
		xmin = as.numeric(factor(unit)) - 0.45,
		xmax = as.numeric(factor(unit)) + 0.45,
		ymin = ymin-0.1,
		ymax = ymax+0.1
	),
	fill = scales::alpha("black", 0.025),
	inherit.aes = FALSE # Prevent the default aesthetic mapping
)

p <- p + geom_point(size = 2)

p
ggsave("preference_distribution.png", plot = p, width = 6, height = 3, dpi = 1200, bg = "white")
```

```{r Comparison of preference rankings and scores, warning=FALSE}
preference_score_model <- translations |> 
	group_by(unit, linguist) |> 
	mutate(CQS_diff = ifelse(is.na(lead(CQS)), CQS - lag(CQS), CQS - lead(CQS))) |> 
	ungroup() |> 
	filter(label == "GPT") |> 
	(\(data) lm(CQS_diff ~ preference, data = data))() |> 
	summary()
preference_score_model$r.squared

translations |> 
	group_by(unit, linguist) |> 
	mutate(CQS_diff = ifelse(is.na(lead(CQS)), CQS - lag(CQS), CQS - lead(CQS))) |> 
	ungroup() |> 
	filter(label == "GPT") |> 
	ggplot(aes(x = CQS_diff, y = preference, fill = linguist)) +
		geom_vline(xintercept = 0, linewidth = 0.5, linetype = "dashed", color = "gray50") +
		geom_boxplot(
			aes(group = preference),
			alpha = 0.2,
			color = "gray50",
			width = 0.4,
			fatten = 1.5,
			linewidth = 0.5,
			show.legend = F
		) +
		stat_boxplot(aes(group = preference), geom = "errorbar", width = 0.2, lwd = 0.5, color = "gray50") +
		# geom_point(shape = 21, size = 2, color = "black", alpha = 0.6) +
		scale_y_continuous(breaks = seq(-2,2), labels = c("Strongly prefer human", "Prefer human", "Neutral", "Prefer GPT", "Strongly prefer GPT")) +
		scale_fill_viridis(discrete = T, option="plasma", labels = c("A", "B", "C")) +
		labs(x = "Difference of Scores: GPT – Human", y = NULL, fill = "Linguist") +
		theme_bw() +
		theme(
			panel.grid.minor.y = element_blank()
		)
ggsave("preference-score_comparison.png", width = 6, height = 3, dpi = 1200, bg = "white")
```

```{r Overall score comparison, message=FALSE, warning=FALSE}
p <- translations |>
	group_by(unit, label) |>
	summarise(unit_mean = mean(CQS)) |>
	group_by(label) |>
	summarise(mean = (mean(unit_mean)), sd = sd(unit_mean)) |>
	ggplot(aes(x = factor(label, levels=c("Human", "GPT")), y = mean, fill = label)) +
		geom_bar(stat = "identity") +
		geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +
		labs(x = NULL, y = "MQM Score", fill = NULL) +
		scale_fill_manual(values = c(col_gpt, col_human)) +
		theme_bw() +
		theme(legend.position = "none", text = element_text(size = 20))
p
ggsave("overall_comparison.png", plot = p, width = 4, height = 4, dpi = 1200)
```

```{r Radar plot, message=FALSE, warning=FALSE}
library(fmsb)
radar_data <- translations |>
	group_by(unit, label) |>
	summarise(
		unit_mean_wrongmedicalterm = mean(EC_wrongmedicalterm),
		unit_mean_mistranslation = mean(EC_mistranslation),
		unit_mean_addition = mean(EC_addition),
		unit_mean_omission = mean(EC_omission),
		unit_mean_grammarspelling = mean(EC_grammarspelling),
		unit_mean_register = mean(EC_register),
		unit_mean_awkward = mean(EC_awkward),
		unit_mean_insensitive = mean(EC_insensitive)
	) |>
	group_by(label) |>
	summarise(
	"Wrong medical term" = mean(unit_mean_wrongmedicalterm), sd_wrongmedicalterm = sd(unit_mean_wrongmedicalterm),
	"Mistranslation     " = mean(unit_mean_mistranslation), sd_mistranslation = sd(unit_mean_mistranslation),
	"Addition    " = mean(unit_mean_addition), sd_addition = sd(unit_mean_addition),
	"Omission    " = mean(unit_mean_omission), sd_omission = sd(unit_mean_omission),
	"Grammar, Spelling, Punctuation" = mean(unit_mean_grammarspelling), sd_grammarspelling = sd(unit_mean_grammarspelling),
	"            Language register" = mean(unit_mean_register), sd_register = sd(unit_mean_register),
	"              Awkward style" = mean(unit_mean_awkward), sd_awkward = sd(unit_mean_awkward),
	"             Culturally insensitive" = mean(unit_mean_insensitive), sd_insensitive = sd(unit_mean_insensitive)
	)
radar_means <- radar_data |> select(-contains("sd")) |> select(-"label")

# Number of errors per translation unit
# radar_table <- as.data.frame(rbind(rep(ceiling(max(radar_means)), ncol(radar_means)), rep(0, ncol(radar_means)), radar_means))

# Proportion of errors (out of total number of errors)
radar_table <- as.data.frame(rbind(rep(round(max(radar_means * 100/sum(radar_means)) / 20) * 20, ncol(radar_means)), rep(0, ncol(radar_means)), radar_means * 100/sum(radar_means)))

rownames(radar_table) <- c("Max", "Min", radar_data$label)
png("radar_chart.png", width = 6400, height = 4800, res = 1200)
labels <- paste0(seq(0,radar_table[1,1],20), "%")
radarchart(radar_table,
		pcol=sapply(c(col_gpt, col_human), adjustcolor, alpha = 0.9),
		pfcol=sapply(c(col_gpt, col_human), adjustcolor, alpha = 0.3),
		plty = 1, cglty=1,
		vlcex = 0.7,
		title="Proportion of All Error Types",
		seg = 3,
		axistype = 0,
		cglcol="grey", axislabcol="grey45",
		calcex = 0.6)
text(
  x = 0.0275,
  y = seq(0,1,0.25)[-1] + 0.05,
  labels = labels,
  adj = c(0, 0.5),
  col = "white",
  cex = 0.625,
  font = 2
)
text(
  x = 0.03,
  y = seq(0,1,0.25)[-1] + 0.05,
  labels = labels,
  adj = c(0, 0.5),
  col = "grey45",
  cex = 0.6
)
legend(x=-2.5, y=-0.05, legend = rownames(radar_table[-c(1,2),]), bty = "n", pch=15, col=c(col_gpt, col_human, col_neutral), cex=1.1, pt.cex=2)
# theme(text = element_text(size = 12))
dev.off()
```

# Statistical testing
```{r Equivalence testing}
# Calculate the GPT and human means and sds
equivtesting <- translations |> 
	select(unit, linguist, label, RQS, CQS) |> 
	pivot_wider(names_from = label, values_from = c(RQS, CQS)) |> 
	mutate(diffRQS = RQS_GPT - RQS_Human, diffCQS = CQS_GPT - CQS_Human) |> 
	group_by(unit) |> 
	summarise(meandiffRQS = mean(diffRQS), meandiffCQS = mean(diffCQS))
# Calculate the CIs
equivtesting_ci <- equivtesting |> 
	summarise(n = length(unique(unit)), 
			  mean_diffRQS = mean(meandiffRQS), sd_diffRQS = sd(meandiffRQS), ciRQS = list(t.test(meandiffRQS, conf.level = 0.95)), 
			  mean_diffCQS = mean(meandiffCQS), sd_diffCQS = sd(meandiffCQS), ciCQS = list(t.test(meandiffCQS, conf.level = 0.95)))
# Print the CIs for RQS and CQS
equivtesting_ci$ciRQS[[1]]
equivtesting_ci$ciCQS[[1]]

# Pick equivtesting_ci$ciRQS or equivtesting_ci$ciCQS
ci <- equivtesting_ci$ciCQS[[1]]
# Define the equivalence region
equivregion <- c(-5,5)
# Plot
p <- ggplot(data = equivtesting_ci, aes(x = mean_diffCQS, y = n)) +
	geom_rect(aes(xmin = equivregion[1], xmax = equivregion[2], ymin = -Inf, ymax = Inf, fill = "Equivalence area"), alpha = 0.1) +
	geom_vline(xintercept=c(-5, 5), linetype="dashed", color="red") +
	geom_vline(xintercept=0) +
	geom_point(size=6, shape=15) +
	geom_errorbarh(aes(xmin = ci$conf.int[1], xmax = ci$conf.int[2]), height = 0.2, linewidth = 1) +
	labs(y = "", x = "Difference: GPT score – Human score", title = "") +
	annotate("text", x = 0, y = equivtesting_ci$n-0.9, hjust = 1.1, label = "Human score > GPT score", color = "gray25", size = 4) +
	annotate("text", x = 0, y = equivtesting_ci$n-0.9, hjust = -0.1, label = "GPT score > Human score", color = "gray25", size = 4) +
	annotate("text", x = equivregion[1], y = equivtesting_ci$n-1, hjust = -0.1, vjust = 1.8, angle = 90, label = "Non-inferiority margin", color = "gray25", size = 3) +
	annotate("text", x = equivregion[2], y = equivtesting_ci$n-1, hjust = -0.2, vjust = -1.0, angle = 90, label = "Superiority margin", color = "gray25", size = 3) +
	expand_limits(x = c(-2.5, 2.5)) + 
	scale_y_continuous(limits = c(equivtesting_ci$n-1,equivtesting_ci$n+1)) +
	scale_fill_manual(values = c("Equivalence area" = "red")) +
	theme_minimal() +
	theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(), panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(), axis.line.x = element_line(color = "black"), plot.title = element_text(hjust=0.5), legend.title = element_blank(), text = element_text(size = 17))
p
ggsave("CQS_forest.png", plot = p, width = 8, height = 2.5, dpi = 600, bg = "white")
```

```{r Misc}
translations_paired <- translations |>
	select(c('unit', 'linguist', 'CQS', 'label')) |>
	pivot_wider(names_from = label, values_from = CQS)
t.test(translations_paired$Human, translations_paired$GPT, paired=TRUE, alternative = "two.sided", conf.level = 0.95)

ec_columns <- names(translations)[startsWith(names(translations), "EC_")]
ttest_results <- lapply(ec_columns, function(col_name) {
  translations_paired <- translations |>
    select(c('unit', 'linguist', all_of(col_name), 'label')) |> 
    group_by(unit, label) |> 
    summarise(mean = mean(!!sym(col_name)), .groups = 'drop') |> 
    pivot_wider(names_from = label, values_from = mean)
  print(paste("****************", col_name, "****************"))
  print(paste("Human:", mean(translations_paired$Human), sd(translations_paired$Human)))
  print(paste("GPT:", mean(translations_paired$GPT), sd(translations_paired$GPT)))
  print(t.test(translations_paired$Human, translations_paired$GPT, paired = TRUE, alternative = "two.sided", conf.level = 0.95))
})
# names(ttest_results) <- ec_columns
# ttest_results

prop.test(c(sum(translations$preference[translations$label == "GPT"] > 0), sum(translations$preference[translations$label == "GPT"] < 0)),
		  c(length(translations$preference[translations$label == "GPT"] > 0), length(translations$preference[translations$label == "GPT"] < 0)))

rateratio(sum(translations$EC_mistranslation[translations$label == "Human"]), sum(translations$EC_mistranslation[translations$label == "GPT"]), length(translations$EC_mistranslation[translations$label == "Human"]), length(translations$EC_mistranslation[translations$label == "GPT"]))
```

```{r Intraclass Correlation Coefficient}
library(psych)
library(lme4)

# load dataframe
df <- read.csv("phase_2_ICC_table.csv")

# create new ID column
df$ID <- paste(df$unit, df$label, sep = "_")


# subset human and GPT translations
df_human <- subset(df, label == "Human")

df_GPT <- subset(df, label == "GPT")


# long to wide
df_wide <- reshape(df, idvar = "ID", timevar = "linguist", direction = "wide", v.names = "RQS")

df_human_wide <- reshape(df_human, idvar = "ID", timevar = "linguist", direction = "wide", v.names = "RQS")

df_GPT_wide <- reshape(df_GPT, idvar = "ID", timevar = "linguist", direction = "wide", v.names = "RQS")


# ICC
# Use ICC2k
icc_results <- ICC(df_wide[, -c(1:3)])
print(icc_results)

icc_human <- ICC(df_human_wide[, -c(1:3)])
print(icc_human)

icc_GPT <- ICC(df_GPT_wide[, -c(1:3)])
print(icc_GPT)
```
